\section{Elements of Financial Statements} 

Activity 1 

For each statement below, tate whether they are True or False. 

\begin{itemize}
    \item A Telephone that is used daily is a current asset.
    \item A machine used to create products is a tangible non-current asset. 
    \item A software licence that lalows a business to use specific software for a period of three years is a tangible non-current asset. 
    \item A truck a business uses to deliver goods to its customers is a tangible non-current asset. 
    \item Goods purchased by a business for resale to its customers are tangible non-current assets. 
\end{itemize}       

Activity 2 

In the following activity, classify the list of items into the elements of financial statements. 

\begin{itemize} 
    \item Shareholder's investment
    \item Computers 
    \item Bookkeeper's annual salary
    \item Warehouse 
    \item Unsold goods
    \item Overdraft with the bank
    \item Cash held at the bank
    \item Sales of goods for cash in the factor shop
\end{itemize} 

Activity 3 

The following activity presents different items that need to be classified under the correct heading in the financial statements. 

\begin{itemize} 
    \item Cost of selling furniture and other costs of running Coltom Co. 
    \item Buildings owned by Coltom Co. 
    \item The amount paid to purchase shares in Coltom Co. 
    \item Amounts Coltom Co. owes for goods and services purchased
    \item The income that Coltom Co. earns from selling furniture to its customers. 
\end{itemize} 


Match each example to the correct financial statement heading below: 

\begin{itemize}
    \item Assets in the Statement of Financial Position 
    \item Liabilities in the Statement of Financial Position 
    \item Equity in the Statement of Financial Position 
    \item Income in the Statement of Profit or Loss 
    \item Expense in the Statement of Profit or Loss
\end{itemize} 


Activity 4 

Classify the following items of expenditure as asset expenditures or expenses: 

\begin{enumerate} 
    \item \$27,000 on a new car. 
    \item \$1,800 road tax incorporated in the car's purchase price in (1) above.
    \item \$10,000 on a second-hand delivery van. 
    \item \$12,000 on refurbishing van in (3) above. 
    \item \$1,000 monthly rental of a vehicle. 
\end{enumerate} 

Activity 5

Rubin owns a business that sells office and computer equipment to corporate customers. His business operates from a warehouse and has a small fleet of delivery 
vehicles. Determine whether teh expenditures made by Rubin's business should be classified as expenses or asset expenditures. 

\begin{enumerate}
    \item Rubin's business has bought some laptops and speakers from its suppliers, which will be sold to its customer. 
    \item Rubin's accountant prepares its financial statements. She orderd a photocopier to make copies of her paperwork.
    \item Rubin's business is expanding, and he has rented a new wareouse. 
    \item Rubin ensures that all the vehicles have valid insurance. Each time he orders a new vehicle, he insures it immediately. 
    \item Goods are delivered to customers using one of the business's delivery vehicles. On the way back to the varehouse, the driver crashhes the side of the vehicle into the fence. The vehicle will need to be fixed to be used again. 
    \item Rubin's business orders a few new vehicles to keep up with customer orders. 
\end{enumerate}

\newpage

Activity 7

Match the equation to the corresponding rearranged accounting equation. 

\begin{table}[h!]
\centering
\begin{tabular}{|l|p{8cm}|l|}
\hline
\textbf{Transaction} & & \textbf{Duality Statement} \\
\hline
Assets & Opening capital + Profit for the period − Drawings & \\
\hline
Closing Net Assets & Closing net assets − Opening net assets + Drawings & \\
\hline
Opening Net Assets & Capital + Liabilities & \\
\hline
Closing Capital & Assets − Capital & \\
\hline
Liabilities & Closing net assets − Profit for the period + Drawings & \\
\hline
Profit for the period & Closing capital & \\
\hline
\end{tabular}
\caption{Transactions and Their Duality Statements}
\end{table}


Activity 8

What is the double entry to record the following transactions? 

\begin{enumerate}
    \item Starts the business by introducing cash.
    \item Purchases some equipment for cash. 
    \item Pays a supplier for some equipment bought on credt. 
    \item Purchases goods for resale on credit. 
    \item Purchases goods for resale for cash. 
    \item Sells goods on credit. 
    \item Sells goods for cash.
\end{enumerate}
